% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode]{acmsmall} % Aptara syntax 

\usepackage{xspace}
\newcommand{\pdht}{{PDHT}\xspace}

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}


% Metadata Information
%\acmVolume{9}
%\acmNumber{4}
%\acmArticle{39}
%\acmYear{2010}
%\acmMonth{3}

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\doi{0000001.0000001}

%ISSN
%\issn{1234-56789}

% Document starts
\begin{document}

% Page heads
%\markboth{G. Zhou et al.}{A Multifrequency MAC Specially Designed for WSN Applications}

% Title portion
\title{Extending a Message Passing Runtime to Support Partitioned, Global Logical Address Spaces}
\author{D. Brian Larkins
\affil{Rhodes College}
James Dinan
\affil{Intel Corporation}}


\begin{abstract}

Partitioned Global Address Space (PGAS) parallel programming models can provide
an efficient mechanism for managing shared data stored across multiple nodes in a
distributed memory system.  However, these models are traditionally directly
addressed and, for applications with loosely-structured or sparse data,
determining the location of a given data element within a PGAS can incur
significant overheads.  When location resolution structures are remote,
applications incur overhead from the network latency of remote lookups.
Further, for large data, caching such structures locally incurs space and
coherence overheads that can limit scaling.
%
We observe that the matching structures used by implementations of the Message
Passing Interface (MPI) establish a separation between incoming data writes and
the location where data will be stored.  In this work, we investigate extending
such structures to add a layer of indirection between incoming data reads
and the location from which data will be read, effectively extending PGAS models
with logical addressing.

\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10003033.10003058.10003064.10011657</concept_id>
<concept_desc>Networks~Network adapters</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010521.10010528</concept_id>
<concept_desc>Computer systems organization~Parallel architectures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011008.10011024.10011028</concept_id>
<concept_desc>Software and its engineering~Data types and structures</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Networks~Network adapters}
\ccsdesc[500]{Computer systems organization~Parallel architectures}
\ccsdesc[500]{Software and its engineering~Data types and structures}


%
% End generated code
%

\keywords{partitioned global address space, parallel data structures, hash tables, hardware acceleration, Portals}

\acmformat{D. Brian Larkins, James Dinan, 2016. Extending a Message Passing Runtime to Support Partitioned, Global Logical Address Spaces}

\begin{bottomstuff}

Author's addresses: D. Brian Larkins, Department of Mathematics and Computer Science,
Rhodes College, Memphis, TN;
James Dinan, Intel Corporation, Hudson, MA.

\end{bottomstuff}

\maketitle

\input{introduction}
\input{background}
\input{implementation}
\input{results}
\input{conclusion}

\cite{cmpi10}

% Acknowledgments
\begin{acks}
This work used the Extreme Science and Engineering Discovery Environment
(XSEDE), which is supported by National Science Foundation grant number
ACI-1053575.
\end{acks}

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{bibdb}
\end{document}
